---
title: "Note-2"
subtitle: "Introduction to DAG"
author: "Haoyue"
date: "`r Sys.Date()`"
output: 
 beamer_presentation:
   latex_engine: xelatex
   slide_level: 3
   theme: Antibes
   colortheme: "beaver"
   fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# The basic settings of this introduction

No $\leftrightarrow$ causal relationships in our analysis

# Roadmap of this note 2

- backdoor path

- collider path

- mediator path

- block the backdoor paths

- backdoor criteria


## Backdoor path

```{r echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdag)
library(dagitty)

backdoor <- dagify(Y ~ D + X,
                   D ~ X,
                   exposure = "D",
                   outcome = "Y")
ggdag_status(backdoor) + theme_dag()
```

---

- X is the variable contributing to the backdoor path

- The existence of X makes the regression $y = \delta d + \varepsilon$ not accurate

- We need to include X to the regression

## Collider path

```{r echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE, paged.print=TRUE}
backdoor <- dagify(X ~ Y + D,
                   Y ~ D,
                   exposure = "D",
                   outcome = "Y")
ggdag_status(backdoor) + theme_dag()
```

---

- collider path is blocked

- we should not include X into the regression

- even if there is X, the regression $y = \delta d + \varepsilon$ is still accurate

- if we condition on the X, the closed collider path is open to create omited variable bias in regression 

## Mediator 

```{r echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE, paged.print=TRUE}
backdoor <- dagify(Y ~ D + X,
                   X ~ D,
                   exposure = "D",
                   outcome = "Y")
ggdag_status(backdoor) + theme_dag()
```

---

the mediator path is usually used in a more complex graph to form backdoor path

## Block backdoor path

- block backdoor path by conditioning on the variable

- it means to add the variable into the regression formula

## Backdoor criteria

- by conditioning on several variables(as few as possible), we can
  + block all backdoor paths
  + we don't open a closed collider path